Verification of the CVM algorithm with a Functional Probabilistic Invariant

### 2025-06-30 paper/sec/formalization.tex
Paper: Additional related work and discussion of probabilistic hoare logic.
Paper: Update equation references.

%TODO: make sure these names are correct!
\begin{note}
In our supplementary material~\cite{CVM_Distinct_Elements-AFP}, the theory \verb|CVM_Abstract_Algorithm| verifies a generalized version of the CVM algorithm, with an abstract subsampling operation that is required to fulfill \cref{i:subsample_condition}.
The specialization happens in the following theories, where \verb|CVM_Original_Algorithm| verifies the original algorithm, and \verb|CVM_New_Unbiased_Algorithm| verifies the new total and unbiased variant.
Note that only \verb|CVM_New_Unbiased_Algorithm| depends on the new library for negatively associated random variables, which we describe in more detail in \cref{sec:formalization_neg_dep}.
The total number of lines required for the verification of the original algorithm is \locnew~lines.
In addition, we actually verified a slight generalization of~\cref{alg:cvm} where the subsampling probability can be any $f \in [\frac{1}{2};e^{-1/12}]$; the original CVM algorithm~\cite{chakraborty2022} is the special case $f=\frac{1}{2}$.
\lipicsEnd\end{note}

>>>>

\begin{note}
In our supplementary material~\cite{CVM_Distinct_Elements-AFP}, the theory \verb|CVM_Abstract_Algorithm| verifies a generalized version of the CVM algorithm, with an abstract subsampling operation that is required to fulfill Inequality~\ref{i:subsample_condition}.
The specialization happens in the following theories, where \verb|CVM_Original_Algorithm| verifies the original algorithm, and \verb|CVM_New_Unbiased_Algorithm| verifies the new total and unbiased variant.
Note that only \verb|CVM_New_Unbiased_Algorithm| depends on the new library for negatively associated random variables, which we describe in more detail in \cref{sec:formalization_neg_dep}.
The total number of lines required for the verification of the original algorithm is \locnew~lines.
In addition, we actually verified a slight generalization of~\cref{alg:cvm} where the subsampling probability can be any $f \in [\frac{1}{2};e^{-1/12}]$; the original CVM algorithm~\cite{chakraborty2022} is the special case $f=\frac{1}{2}$.
\lipicsEnd\end{note}

### 2025-06-30 paper/sec/invariants.tex
Paper: References to supplementary material.   - Instead of referring to directories in the anonymized zip file, we refer to the AFP entry and github repo.   - Improve related work section regarding distinct-elements algorithms from Bar-Yossed and Alon et al.   - Add license files (use BSD consistent with the license of the AFP entries.)   - Add readme for the CVM_Transform session which is going to be references by swhids.
Paper: Update equation references.

Let us see how we can verify \cref{eq:indicator_eq} more formally.
For that, we model the state of the randomized algorithm as a pair $(\chi,p)$ and we write $\chi$ and $p$ for the random variables projecting their respective components from the distribution of the state of the algorithm.
%For that, let us describe the states of the algorithm using pairs composed of a subset of $A$ representing the buffer $\chi$ and a real number representing $p$.
%Note that, this means $p$ and $\chi$ are random variables, with respect to distributions of the state of the algorithm.
We will refer to parts of each loop iteration in \cref{alg:cvm} as $\mathrm{step}_1$ (resp.~$\mathrm{step}_2$) for Lines 3--7 (resp.~Lines 8--10).
The final distribution of the algorithm is the distribution resulting from the sequential composition of alternating steps over the stream:
\[
  \mathrm{init} \, \isa{\isasymbind}\; \mathrm{step}_1\, a_1 \; \isa{\isasymbind}\; \mathrm{step}_2 \; \isa{\isasymbind}\; \mathrm{step}_1 \, a_2 \; \isa{\isasymbind}\; \cdots \isa{\isasymbind}\; \mathrm{step}_1\, a_l \; \isa{\isasymbind}\; \mathrm{step}_2
\]
where we parameterize $\mathrm{step}_1$ with the stream element that it processes.
The term $\mathrm{init}$ represents the initial state, i.e., $\mathrm{init} = \mathrm{\bf return}\,(\{\},1)$.
It is easy to show by induction over the sequence of steps, we have $0 < p \leq 1$ and $\chi \subseteq A$ for all possible states of the algorithm.

Let us verify that~\cref{eq:indicator_eq} is preserved as an invariant over all steps.
To verify that $\mathrm{step}_1\, a$ preserves \cref{eq:indicator_eq}, we assume some probability space of states $\Omega$ fulfills \cref{eq:indicator_eq} and we would like to show that it is still true for $\Omega \; \isa{\isasymbind}\; \mathrm{step}_1\, a$. By~\cref{eq:integral_bind},
\begin{equation}\label{eq:step_1_exp}
  \expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step}_1\, a} [ p^{-1} X_s ] =
    \integral{\Omega}{\integral{\Ber(p)}{p^{-1} \indicat\left(s \in (\ift{\tau}{\chi \cup \{a\}}{\chi-\{a\}} )\right)}{\tau}}{\sigma} \textrm{.}
    % \int_\Omega \int_{\Ber(p)} p^{-1} \indicat\left(s \in (\ift{\tau}{\chi \cup \{a\}}{\chi-\{a\}} )\right) \, d \tau d \sigma \textrm{.}
\end{equation}
Note that we write $p$ or $\chi$ even though we should actually write $p(\sigma)$ or $\chi(\sigma)$, i.e., we remember that these implicitly depend on $\sigma$.
To see that the right-hand side is equal to $1$, it is useful to consider cases on whether $a=s$.
When $a = s$, the right-hand-side is equal to $1$ by definition of the Bernoulli distribution (since $p \in (0;1]$).
When $a \not= s$, it follows from the induction hypothesis on $\Omega$; in particular, the term in the inner integral is constant with respect to $\tau$.

The same invariant-based argument is possible for $\mathrm{step}_2$.
Let us assume $\Omega$ is a probability space of states fulfilling \cref{eq:indicator_eq}.
Then by~\cref{eq:integral_bind}, $\expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} [p^{-1}X_s]$ equals
\begin{equation}\label{eq:step_2_exp}
%  \expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} \left[\frac{X_s}{p}\right]
    \integral{\Omega}
      {\left(\ift{|\chi|=n}{\left(\integral{\mathrm{subsample}(\chi)}{\frac{\indicat(s \in \tau)}{p/2}}{\tau}\right)}{\frac{\indicat(s \in \chi)}{p}} \right)}
      {\sigma}
    \textrm{.}
\end{equation}
Note that the \emph{true} and \emph{false} cases of the inner \textbf{if-then-else} both evaluate to the same value: $p^{-1} \indicat(s \in \chi)$.
If $s \notin \chi$ both sides of the equation are $0$, because the subsampling operation returns a subset of $\chi$.
If $s \in \chi$ the probability that the element gets subsampled is $1/2$, so we arrive again at $\frac{1/2}{p/2} = p^{-1} \indicat(s \in \chi)$.
Hence: $\expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} [p^{-1} X_s] = \expect_\Omega [p^{-1} X_s] = 1$.
This completes the invariance proof for~\cref{eq:indicator_eq}.

\subsection{Deriving a Functional Probabilistic Invariant}
With \cref{eq:indicator_eq} established, it is straightforward to show that the expected value of the output estimate $p^{-1} |\chi|$ for the modified algorithm (without Line 11) is equal to the desired cardinality $|A|$.
However, recall that we are interested in verifying the estimate's PAC guarantee.
A typical approach to establishing such a guarantee is to use Chernoff bounds which provide exponential tail bounds (i.e., concentration bounds) for the deviation of sums of independent random variables from their mean.
However, these are not directly useful in the CVM algorithm because the key random variables, e.g., $p^{-1} X_s$ for $s \in A$, are dependent.

@@

In our case, we are interested in estimating the moment generating function of the random variable $p^{-1} |\chi|$ for the CVM algorithm:
\[
  \expect [\exp( t p^{-1} |\chi| )] = \expect \left[ \prod_{s \in A} h(p^{-1} X_s) \right]
\]
for $h(x) = \exp(tx)$.
At this point, it is tempting to see whether the proof for \cref{eq:indicator_eq} can be generalized to establish bounds for the above.
Indeed, we managed to establish the following generalized result:
\begin{align}
  \expect \left[ \prod_{s \in A} h(p^{-1} X_s) \right] \leq h(1)^{|A|} \label[ineq]{i:func_invariant}
\end{align}
for every non-negative concave function $h : \Rnonneg \rightarrow \Rnonneg$.
We call this a \emph{functional probabilistic invariant} because one can establish it for all valid choices of $h$ with a single induction over steps of the randomized algorithm.

While the exponential function in $M(t)$ is convex, we can instead try to derive tail bounds for the random variable $\indicat(p \geq q) p^{-1} |\chi|$, for some fixed constant $q > 0$.
This leads to a similar invariant inequality:
\begin{align}
  \expect \left[ \prod_{s \in A} I(p \geq q) h(p^{-1} X_s) \right] \leq h(1)^{|A|} \label[ineq]{i:func_invariant_capped}
\end{align}
with the new condition that $h$ needs to be non-negative and concave only on $[0;q^{-1}]$.
This then allows us to approximate the exponential function from above with an affine function $h$ on the range $[0;q^{-1}]$, which yields tail bounds for $p^{-1} |\chi|$ under the condition $p \geq q$. As an example, the upper tail bound can be derived as follows:
%TODO: should we use exp(...) everywhere instead of mixing with e^ ? EK: That was the previous verison, main issue is the chain looks very lopsided, with almost all of the terms being on the right edge. (It looked too ugly.)--- OK
% The convention is now, if exp is applied to a random variable we use the long notation and otherwise the power notation, which might actually be (a bit) helpful.
\begin{eqnarray*}
  \prob( p^{-1} |\chi| \geq (1+\varepsilon) |A| \wedge p \geq q) & \leq & % Is not equal when |A|=0.
  \prob( \indicat(p \geq q) p^{-1} |\chi| \geq (1+\varepsilon) |A|) \\
  & \underset{\textrm{Markov}}{\leq} & e^{-t(1+\varepsilon) |A|} \expect \left[ \prod_{s \in A} \indicat(p \geq q) \exp( t p^{-1} X_s) \right] \\
  & \leq & e^{-t(1+\varepsilon) |A|} \expect \left[ \prod_{s \in A} \indicat(p \geq q) h(p^{-1} X_s) \right] \\
  & \underset{\textrm{Ineq.~\ref{i:func_invariant_capped}}}{\leq} & e^{-t(1+\varepsilon) |A|} h(1)^{|A|} \\
  & \underset{\textrm{Calculus}}{\leq} & e^{-n\varepsilon^2 / 12}
\end{eqnarray*}
where we choose $h(x) = 1+qx (e^{t/q}-1)$.
Note that $h$ is affine and it can be easily checked%
\footnote{Because the exponential function is convex and $h$ is affine, we only have to check the end points: $0, q^{-1}$.}
that it is an upper approximation of $\exp(tx)$ for $x \in [0;q^{-1}]$.
For the last step, we have to find the $t$ that produces the required bound.%
\footnote{We use $t = q \ln(1+\varepsilon)$ which is not the real optimum, but better for algebraic evaluation.}
To use these bounds, we also have to separately estimate $\prob(p < q)$.
For that, we use a similar strategy, as in the original proof by Chakraborty et al.~\cite{chakraborty2023}, with $q = \frac{n}{4 |A|}$.
%The formalization accompanying this work~\todo{[cite]} contains a detailed informal step-by-step proof using our approach in its appendix.
The formalization in the supplementary material contains a detailed informal step-by-step proof using our approach in its appendix.
Besides the use of \cref{eq:integral_bind} and the Cram\'er--Chernoff method, the steps are elementary.

>>>>

Let us see how we can verify Equation~\ref{eq:indicator_eq} more formally.
For that, we model the state of the randomized algorithm as a pair $(\chi,p)$ and we write $\chi$ and $p$ for the random variables projecting their respective components from the distribution of the state of the algorithm.
%For that, let us describe the states of the algorithm using pairs composed of a subset of $A$ representing the buffer $\chi$ and a real number representing $p$.
%Note that, this means $p$ and $\chi$ are random variables, with respect to distributions of the state of the algorithm.
We will refer to parts of each loop iteration in \cref{alg:cvm} as $\mathrm{step}_1$ (resp.~$\mathrm{step}_2$) for Lines 3--7 (resp.~Lines 8--10).
The final distribution of the algorithm is the distribution resulting from the sequential composition of alternating steps over the stream:
\[
  \mathrm{init} \, \isa{\isasymbind}\; \mathrm{step}_1\, a_1 \; \isa{\isasymbind}\; \mathrm{step}_2 \; \isa{\isasymbind}\; \mathrm{step}_1 \, a_2 \; \isa{\isasymbind}\; \cdots \isa{\isasymbind}\; \mathrm{step}_1\, a_l \; \isa{\isasymbind}\; \mathrm{step}_2
\]
where we parameterize $\mathrm{step}_1$ with the stream element that it processes.
The term $\mathrm{init}$ represents the initial state, i.e., $\mathrm{init} = \mathrm{\bf return}\,(\{\},1)$.
It is easy to show by induction over the sequence of steps, we have $0 < p \leq 1$ and $\chi \subseteq A$ for all possible states of the algorithm.

Let us verify that Equation~\ref{eq:indicator_eq} is preserved as an invariant over all steps.
To verify that $\mathrm{step}_1\, a$ preserves Equation~\ref{eq:indicator_eq}, we assume some probability space of states $\Omega$ fulfills Equation~\ref{eq:indicator_eq} and we would like to show that it is still true for $\Omega \; \isa{\isasymbind}\; \mathrm{step}_1\, a$. By Equation~\ref{eq:integral_bind},
\begin{equation}\label{eq:step_1_exp}
  \expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step}_1\, a} [ p^{-1} X_s ] =
    \integral{\Omega}{\integral{\Ber(p)}{p^{-1} \indicat\left(s \in (\ift{\tau}{\chi \cup \{a\}}{\chi-\{a\}} )\right)}{\tau}}{\sigma} \textrm{.}
    % \int_\Omega \int_{\Ber(p)} p^{-1} \indicat\left(s \in (\ift{\tau}{\chi \cup \{a\}}{\chi-\{a\}} )\right) \, d \tau d \sigma \textrm{.}
\end{equation}
Note that we write $p$ or $\chi$ even though we should actually write $p(\sigma)$ or $\chi(\sigma)$, i.e., we remember that these implicitly depend on $\sigma$.
To see that the right-hand side is equal to $1$, it is useful to consider cases on whether $a=s$.
When $a = s$, the right-hand-side is equal to $1$ by definition of the Bernoulli distribution (since $p \in (0;1]$).
When $a \not= s$, it follows from the induction hypothesis on $\Omega$; in particular, the term in the inner integral is constant with respect to $\tau$.

The same invariant-based argument is possible for $\mathrm{step}_2$.
Let us assume $\Omega$ is a probability space of states fulfilling Equation~\ref{eq:indicator_eq}.
Then by Equation~\ref{eq:integral_bind}, $\expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} [p^{-1}X_s]$ equals
\begin{equation}\label{eq:step_2_exp}
%  \expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} \left[\frac{X_s}{p}\right]
    \integral{\Omega}
      {\left(\ift{|\chi|=n}{\left(\integral{\mathrm{subsample}(\chi)}{\frac{\indicat(s \in \tau)}{p/2}}{\tau}\right)}{\frac{\indicat(s \in \chi)}{p}} \right)}
      {\sigma}
    \textrm{.}
\end{equation}
Note that the \emph{true} and \emph{false} cases of the inner \textbf{if-then-else} both evaluate to the same value: $p^{-1} \indicat(s \in \chi)$.
If $s \notin \chi$ both sides of the equation are $0$, because the subsampling operation returns a subset of $\chi$.
If $s \in \chi$ the probability that the element gets subsampled is $1/2$, so we arrive again at $\frac{1/2}{p/2} = p^{-1} \indicat(s \in \chi)$.
Hence: $\expect_{\Omega \; \isa{\isasymbind}\; \mathrm{step_2}} [p^{-1} X_s] = \expect_\Omega [p^{-1} X_s] = 1$.
This completes the invariance proof for Equation~\ref{eq:indicator_eq}.

\subsection{Deriving a Functional Probabilistic Invariant}
With Equation~\ref{eq:indicator_eq} established, it is straightforward to show that the expected value of the output estimate $p^{-1} |\chi|$ for the modified algorithm (without Line 11) is equal to the desired cardinality $|A|$.
However, recall that we are interested in verifying the estimate's PAC guarantee.
A typical approach to establishing such a guarantee is to use Chernoff bounds which provide exponential tail bounds (i.e., concentration bounds) for the deviation of sums of independent random variables from their mean.
However, these are not directly useful in the CVM algorithm because the key random variables, e.g., $p^{-1} X_s$ for $s \in A$, are dependent.

@@

In our case, we are interested in estimating the moment generating function of the random variable $p^{-1} |\chi|$ for the CVM algorithm:
\[
  \expect [\exp( t p^{-1} |\chi| )] = \expect \left[ \prod_{s \in A} h(p^{-1} X_s) \right]
\]
for $h(x) = \exp(tx)$.
At this point, it is tempting to see whether the proof for Equation~\ref{eq:indicator_eq} can be generalized to establish bounds for the above.
Indeed, we managed to establish the following generalized result:
\begin{align}
  \expect \left[ \prod_{s \in A} h(p^{-1} X_s) \right] \leq h(1)^{|A|} \label[ineq]{i:func_invariant}
\end{align}
for every non-negative concave function $h : \Rnonneg \rightarrow \Rnonneg$.
We call this a \emph{functional probabilistic invariant} because one can establish it for all valid choices of $h$ with a single induction over steps of the randomized algorithm.

While the exponential function in $M(t)$ is convex, we can instead try to derive tail bounds for the random variable $\indicat(p \geq q) p^{-1} |\chi|$, for some fixed constant $q > 0$.
This leads to a similar invariant inequality:
\begin{align}
  \expect \left[ \prod_{s \in A} I(p \geq q) h(p^{-1} X_s) \right] \leq h(1)^{|A|} \label[ineq]{i:func_invariant_capped}
\end{align}
with the new condition that $h$ needs to be non-negative and concave only on $[0;q^{-1}]$.
This then allows us to approximate the exponential function from above with an affine function $h$ on the range $[0;q^{-1}]$, which yields tail bounds for $p^{-1} |\chi|$ under the condition $p \geq q$. As an example, the upper tail bound can be derived as follows:
%TODO: should we use exp(...) everywhere instead of mixing with e^ ? EK: That was the previous verison, main issue is the chain looks very lopsided, with almost all of the terms being on the right edge. (It looked too ugly.)--- OK
% The convention is now, if exp is applied to a random variable we use the long notation and otherwise the power notation, which might actually be (a bit) helpful.
\begin{eqnarray*}
  \prob( p^{-1} |\chi| \geq (1+\varepsilon) |A| \wedge p \geq q) & \leq & % Is not equal when |A|=0.
  \prob( \indicat(p \geq q) p^{-1} |\chi| \geq (1+\varepsilon) |A|) \\
  & \underset{\textrm{Markov}}{\leq} & e^{-t(1+\varepsilon) |A|} \expect \left[ \prod_{s \in A} \indicat(p \geq q) \exp( t p^{-1} X_s) \right] \\
  & \leq & e^{-t(1+\varepsilon) |A|} \expect \left[ \prod_{s \in A} \indicat(p \geq q) h(p^{-1} X_s) \right] \\
  & \underset{\textrm{Ineq.~\ref{i:func_invariant_capped}}}{\leq} & e^{-t(1+\varepsilon) |A|} h(1)^{|A|} \\
  & \underset{\textrm{Calculus}}{\leq} & e^{-n\varepsilon^2 / 12}
\end{eqnarray*}
where we choose $h(x) = 1+qx (e^{t/q}-1)$.
Note that $h$ is affine and it can be easily checked%
\footnote{Because the exponential function is convex and $h$ is affine, we only have to check the end points: $0, q^{-1}$.}
that it is an upper approximation of $\exp(tx)$ for $x \in [0;q^{-1}]$.
For the last step, we have to find the $t$ that produces the required bound.%
\footnote{We use $t = q \ln(1+\varepsilon)$ which is not the real optimum, but better for algebraic evaluation.}
To use these bounds, we also have to separately estimate $\prob(p < q)$.
For that, we use a similar strategy, as in the original proof by Chakraborty et al.~\cite{chakraborty2023}, with $q = \frac{n}{4 |A|}$.
%The formalization accompanying this work~\todo{[cite]} contains a detailed informal step-by-step proof using our approach in its appendix.
The formalization in the supplementary material~\cite{CVM_Distinct_Elements-AFP} contains a detailed informal step-by-step proof using our approach in its appendix.
Besides the use of Equation~\ref{eq:integral_bind} and the Cram\'er--Chernoff method, the steps are elementary.

### 2025-06-30 paper/sec/negdep.tex
Paper: Update equation references.

When we look at the subsampling step of~\cref{alg:cvm}, our invariant~\cref{i:func_invariant} imposes the following condition on the subsampling operation.
It should be noted that the condition becomes apparent while establishing \cref{i:func_invariant} using similar but more general steps as described in Equations~\ref{eq:step_1_exp} and \ref{eq:step_2_exp}.%
\footnote{A step-by-step proof of the derivation is also available in the appendix of the formalization~\cite[Appendix~A]{CVM_Distinct_Elements-AFP}.}
\begin{equation}\label[ineq]{i:subsample_condition}
  \integral{\mathrm{subsample}(\chi)}{\prod_{s \in S} g(\indicat(s \in \tau))}{\tau} \leq \prod_{s \in S} \expect_{\Ber(f)} [g]
\end{equation}
for all non-negative functions $g$ and any $S \subseteq \chi$, where $f$ is the probability of retaining each element in the subsampling step of the algorithm.
Any subsampling step that satisfies this functional inequality can be used while still preserving~\cref{i:func_invariant} for the algorithm.

Motivated by this observation, our new variant is shown in \cref{alg:cvm_new}.
For the subsampling step, instead of keeping each element of $\chi$ with probability $\frac{1}{2}$, we pick a uniform random $nf$-subset of $\chi$, where $\frac{1}{2} \leq f < 1$ and $nf$ is an integer.
For example, it is possible to choose $f = \frac{n-1}{n}$, i.e., discarding a random element from $\chi$ in the subsampling step.
Since this new subsampling step always reduces the size of $\chi$, the variant is \emph{total} (never returns $\bot$).
The invariant-based approach allows us to show that the algorithm is probably-approximately correct and also \emph{unbiased}, i.e., the expectation of the result is exactly $|A|$.
These depend crucially on establishing~\cref{i:subsample_condition} for the new subsampler, for which we need a new concept.

>>>>

When we look at the subsampling step of~\cref{alg:cvm}, our invariant (Inequality~\ref{i:func_invariant}) imposes the following condition on the subsampling operation.
It should be noted that the condition becomes apparent while establishing Inequality~\ref{i:func_invariant} using similar but more general steps as described in Equations~\ref{eq:step_1_exp} and \ref{eq:step_2_exp}.%
\footnote{A step-by-step proof of the derivation is also available in the appendix of the formalization~\cite[Appendix~A]{CVM_Distinct_Elements-AFP}.}
\begin{equation}\label[ineq]{i:subsample_condition}
  \integral{\mathrm{subsample}(\chi)}{\prod_{s \in S} g(\indicat(s \in \tau))}{\tau} \leq \prod_{s \in S} \expect_{\Ber(f)} [g]
\end{equation}
for all non-negative functions $g$ and any $S \subseteq \chi$, where $f$ is the probability of retaining each element in the subsampling step of the algorithm.
Any subsampling step that satisfies this functional inequality can be used while still preserving Inequality~\ref{i:func_invariant} for the algorithm.

Motivated by this observation, our new variant is shown in \cref{alg:cvm_new}.
For the subsampling step, instead of keeping each element of $\chi$ with probability $\frac{1}{2}$, we pick a uniform random $nf$-subset of $\chi$, where $\frac{1}{2} \leq f < 1$ and $nf$ is an integer.
For example, it is possible to choose $f = \frac{n-1}{n}$, i.e., discarding a random element from $\chi$ in the subsampling step.
Since this new subsampling step always reduces the size of $\chi$, the variant is \emph{total} (never returns $\bot$).
The invariant-based approach allows us to show that the algorithm is probably-approximately correct and also \emph{unbiased}, i.e., the expectation of the result is exactly $|A|$.
These depend crucially on establishing Inequality~\ref{i:subsample_condition} for the new subsampler, for which we need a new concept.
